


```{r setup, include=FALSE, cache=FALSE}
#Set root directory to R project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```


```{css, echo=FALSE}
# set max height of code chunks using the following css styling

pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 300px;
}
```

## Setup

### Libraries

```{r}
library(rgdal)
library(dplyr)
library(tidyr)
library(sf)
library(tidyverse)
library(raster)
```


### Load unclipped data

```{r eval=FALSE, echo=FALSE}
# unclipped data

####### pinch-points
#load("1_data/manual/pp_all_RoG.rdata")
#pp_all_clus_RoG$group<- paste(pp_all_clus_RoG$ID, pp_all_clus_RoG$Reach, sep = "_")
  
####### covariates

#UPLVI
UPLVI<-st_read("1_data/external/2018_uPLVI/COE_2018_uPLVI_RES_UPDATE_DRAFTv3.shp")

#picnic
picnic<-st_read("1_data/external/Parks_Infrastructure/Copy_of_PICNIC_SITES.shp")
st_crs(picnic)
picnic<-st_transform(picnic, crs = st_crs(3776))

#playground
playgrounds<-st_read("1_data/external/Parks_Infrastructure/Copy_of_Playgrounds.shp")
st_crs(playgrounds)
playgrounds<-st_transform(playgrounds, crs = st_crs(3776))

#sports fields
sports_fields<-st_read("1_data/external/Parks_Infrastructure/Copy_of_SPORTS_FIELDS.shp")
st_crs(sports_fields)
sports_fields<-st_transform(sports_fields, crs = st_crs(3776))

# there are some invalid geometries in this spatial layer.
# identify invalid geometries:
valid<-st_is_valid(sports_fields)

#fix by by buffering geometries:
sports_fields<-st_buffer(sports_fields[!is.na(valid),], 0.0)




#slope
slope<-raster("1_data/external/Edmonton_slope_aspect_in_degrees/slope")
slope<-projectRaster(slope, crs=crs(picnic))



#circuitscape raster
cur_map_sum<-raster("1_data/external/Circuitscape/Map/Coyote_Summer_All_curmap.tif")
cur_map_sum<-projectRaster(cur_map_sum, crs=crs(picnic))

cur_map_win<-raster("1_data/external/Circuitscape/Map/Coyote_Winter_All_curmap.tif")
cur_map_win<-projectRaster(cur_map_win, crs=crs(picnic))

# #Canopy polygons
# can_path<-"1_data/external/Canopy_Cover_2020/PolygonData.gdb"
# # # List all feature classes in a file geodatabase
# # subset(ogrDrivers(), grepl("GDB", name))
# # fc_list <- ogrListLayers(can_path)
# # print(fc_list)
# can_poly <- readOGR(dsn=can_path,layer="All_Polygons")



# load ribbon of green
RoG_Reaches<-st_read("1_data/external/Ribbon_of_Green_Study_Area_Reaches/Reach_Boundaries/RoG_Reaches_V1_210301/RoG_Reaches_V1_210301.shp")


# check crs
st_crs(RoG_Reaches)
# NAD83_3TM_114_Longitude_Meter_Province_of_Alberta_Canada
RoG_Reaches<-st_transform(RoG_Reaches, crs = st_crs(3776))

```

load pp clusters

```{r}
load("1_data/manual/pp_all_RoG.rdata")
```


## Metrics in pinch_point

### Calculate proportion of cover

#### Intersect polygon layers with pinch points

```{r}
# list objects to individually intersect by the pinchpoints
#x<-list(UPLVI, can_poly, picnic, playgrounds, sports_fields)
x<-list(UPLVI, picnic, playgrounds, sports_fields)
# names(x)<-c("UPLVI_int", "can_poly_int", "picnic_int", "playgrounds_int", "sports_fields_int")
names(x)<-c("UPLVI_int", "picnic_int", "playgrounds_int", "sports_fields_int")


# calculate the area of the buffered pinchpoints
pp_all_clus_RoG$in_area<-st_area(pp_all_clus_RoG)


#apply the following to all the layers
nx<-lapply(x, function(x) 
  {
  a<-st_intersection(x, pp_all_clus_RoG) 
  cbind(a, "int_area"=st_area(a)) #calculate area of new clipped polygons
  })

#save new intersection objects
lapply(names(nx),function(x)
  { assign(x,nx[[x]])
  save(list=x, file=paste("3_pipeline/tmp/", x, "_in.rdata", sep = ""))
  })

# put intersection objects into the global environment
list2env(nx, .GlobalEnv)
```

#### Proportion  of cover

##### From UPLVI {-}

```{r eval=FALSE}
load("3_pipeline/tmp/UPLVI_int_in.rData")
UPLVI_int$group<- paste(UPLVI_int$ID, UPLVI_int$Reach, sep = "_")


#UPLVI_int$ID <- paste(UPLVI_int$ID, UPLVI_int$Reach, sep = "_")


#UPLVI_int$group<- paste(UPLVI_int$ID, UPLVI_int$Reach, sep = "_")

# create a new dataframe with only necessary columns

TYPE<-as.data.frame(UPLVI_int[c(1,61:62, 4:7, 23:26, 42:45, 63:67)])

#seperate STYPPE ranks into seperate dataframes
a<-TYPE[c(1:3, 4:7, 16:20)]
b<-TYPE[c(1:3, 8:11, 16:20)]
c<-TYPE[c(1:3, 12:15, 16:20)]

# rename columns so they are constant
b<-setNames(b, names(a))
c<-setNames(c, names(a))

#rbind dataframes together
abc<-rbind(a, b, c)

# create a column with just rank number of STYPES
s_rank<-gather(TYPE, "STYPE_rank", "STYPE", c(4,8,12))

# add the rank column
abcf<-cbind(abc, s_rank$STYPE_rank)

#reorder columns
#abcf<-abcf[c(1:3, 8, 4:7)]

# rename columns
colnames(abcf)[c(4:7, 13)]<-c("PRIMECLAS", "LANDCLAS", "STYPE", "STYPEPER", "s_rank")

# approximate area of each STYPE class/ polygon by multipltying the percentage by polygon area. Creates a new column
# calculate proportion of each polygon covered by STYPE class

# here is a way to loop through a bunch of column
v<- abcf[4:6] #select the columns that you want to use for landcover data
vnames <- colnames(v)
a <- list()
j = 1
for (i in v) {
  a[[j]] <- abcf %>%
    group_by(POLY_NUM, group,{{i}})%>%
    mutate(pp_area=mean(area))%>%
    mutate(s_area=as.numeric((STYPEPER*.1)*int_area))%>%
    group_by(group, {{i}})%>%
    mutate(t_area=sum(s_area))%>%
    dplyr::select(-s_rank, -STYPEPER, -s_area, -int_area, -POLY_NUM,  -LANDCLAS,-STYPE, -PRIMECLAS)%>%
    distinct()%>%
    drop_na('<chr>')%>%
    mutate(proportion=as.numeric(t_area/in_area))%>%
    replace_na(list(proportion=0))  %>%
    mutate(metric=vnames[j])%>%
    rename(c("class"="<chr>"))%>%
  dplyr::select(ID, group, season, Reach, in_area, metric, class, proportion)%>%
  arrange(ID, group, season, Reach, in_area, metric, class, proportion)
  names(a)[j]=vnames[j] #name list item
  colnames(a[[j]])[7]<- "class" # name variable column
  a[[j]]$vnames = vnames[j]
  j = j + 1
}
a

# test<-a$PRIMECLAS
# nrow(distinct(as.data.frame(test$group)))


UPLVI_class_cov_all_in<-a

save(UPLVI_class_cov_all_in, file="3_pipeline/store/UPLVI_class_cov_all_in.rData")


```

##### From Canopy polygons {-}

```{r}
can_poly_int<-st_read("3_pipeline/tmp/pp_all_clus_RoG_canopy_1.shp")
#test<-distinct(can_poly_int)

# convert geometry to polygons so can_poly_int plots properly
#can_poly_int<-st_collection_extract(can_poly_int, "POLYGON")

can_poly_int$group<- paste(can_poly_int$ID, can_poly_int$Reach, sep = "_")
can_poly_int$int_area<-st_area(can_poly_int)

x<-can_poly_int


df<-as.data.frame(pp_all_clus_RoG$group)
df<-rename(df, "group"="pp_all_clus_RoG$group")
x<-full_join(x, df, by="group")
#nrow(distinct(as.data.frame(x$group)))


xSum<-as.data.frame(x) %>%
  group_by(ID, group, season, Reach, Class)%>%
  summarize(pp_area=mean(area), int_mean=mean(int_area))%>%                                                                 mutate(proportion=as.numeric(int_mean/sum(int_mean)))%>%
  replace_na(list(proportion=0))  %>%        
  mutate(metric="Vegetation_type")  %>%    
  select(ID, group, season, Reach, pp_area, metric, Class, proportion)%>%
  arrange(ID, group, season, Reach, pp_area, metric, Class, proportion)%>%
  rename("class"="Class") 
  

can_poly_cov_all<-xSum

save(can_poly_cov_all, file="3_pipeline/store/can_poly_cov_all_in.rData")
```


#### Proportion of pinch point within set distance to a road


#### Total length of road in pinch-point buffer

```{r}
########## just in pp############
road_ws<-st_read("1_data/external/roads_data/Road_Width_and_Speed/Road_Width_and_Speed.shp")
st_crs(road_ws)
road_ws<-st_transform(road_ws, crs = st_crs(3776))


tot_road_in<-pp_all_clus_RoG

int_buf = st_intersection(road_ws, tot_road_in)

# plot(polys_buf$geom)
# plot(int_buf$geom, lwd = 2, add = TRUE)

#Get the total length of road within each polygon. (polygons with no lines are missing from list)
road_tot<-tapply(st_length(int_buf), int_buf$ID.1,sum)

road_tot<-as.data.frame(road_tot)
road_tot<-road_tot%>%
  rownames_to_column()%>%
  rename("ID"="rowname")

tot_road_in<-left_join(tot_road_in, road_tot)


#plot(polys_buf[,"road_tot_buf"])
save(tot_road_in, file = "3_pipeline/tmp/tot_road_in.rData")
```





#### Proportion of pinch point within set distance to a trail

```{r}
trails<-st_read("1_data/external/Trails/Copy_of_TRAILS.shp")
st_crs(trails)
trails<-st_transform(trails, crs = st_crs(3776))

```

#### Total length of trail in pinch-points buf

```{r}
########## just in buffer ############
tot_trail_in_buf<-pp_all_clus_RoG

int_buf = st_intersection(trails, tot_trail_in_buf)

# plot(polys_buf$geom)
# plot(int_buf$geom, lwd = 2, add = TRUE)

#Get the total length of road within each polygon. (polygons with no lines are missing from list)
trail_tot<-tapply(st_length(int_buf), int_buf$ID.1,sum)

trail_tot<-as.data.frame(trail_tot)
trail_tot<-trail_tot%>%
  rownames_to_column()%>%
  rename("ID"="rowname")

tot_trail_in<-left_join(tot_trail_in_buf, trail_tot)

# tot_trail_in_buf$trail_tot_buf = rep(0,nrow(polys_buf))
# tot_trail_in_buf$trail_tot_buf[match(names(trail_tot_buf),polys_buf$ID)] = trail_tot_buf

#plot(polys_buf[,"road_tot_buf"])
save(tot_trail_in, file = "3_pipeline/tmp/tot_trail_in.rData")
```







### Raster extract

#### Slope

```{r eval=FALSE}
load("3_pipeline/store/pp_all_clus_RoG.rData")

pp_slope_mean<-raster::extract(slope,  pp_all_clus_RoG, fun=mean, na.rm=TRUE)

colnames(pp_slope_mean)<-"slope_mean"

```




#### Merge raster data into single data frame
```{r}
load("3_pipeline/tmp/pp_ext_r.rData")
pp_cur<-pp_ext_r[-c(9)]

pp_ext_r_in<-cbind(pp_cur, pp_slope_mean)

save(pp_ext_r_in, file="3_pipeline/tmp/pp_ext_r_in.rData")
```



### Combine metrics into a single data frame

### Load data
```{r}
load("1_data/manual/pp_all_RoG.rdata")
pp_all_clus_RoG$group<- paste(pp_all_clus_RoG$ID, pp_all_clus_RoG$Reach, sep = "_")
load("3_pipeline/tmp/pp_ext_r_cent.rData")
pp_cent<-pp_ext_r_cent[-c(11)]
load("3_pipeline/tmp/pp_ext_r_in.rData")
pp_ext_r_cent_in<-left_join(as.data.frame(pp_ext_r_in), pp_cent)



load("3_pipeline/store/prop_Trail_all.rData")



load("3_pipeline/store/prop_Road_all.rData")
#load("3_pipeline/store/can_poly_cov_all.rData")
load("3_pipeline/store/UPLVI_class_cov_all_in.rData")
load("3_pipeline/tmp/tot_trail_in.rData")
tot_trail_in$group<- paste(tot_trail_in$ID, tot_trail_in$Reach, sep = "_")

load("3_pipeline/tmp/tot_road_in.rData")
tot_road_in$group<- paste(tot_road_in$ID, tot_road_in$Reach, sep = "_")

load("3_pipeline/tmp/pp_ext_r_in.rData")


load("1_data/manual/pp_all_RoG.rdata")
```



#### Reformat and join {-}

##### wide format (for PCA)
```{r}
library(units)
pp_ext_r_cent_in<- pp_ext_r_cent_in%>% 
  rename("pp_area"="area")

pp_ext_r_cent_in<-as.data.frame(pp_ext_r_cent_in)
pp_ext_r_cent_in<-pp_ext_r_cent_in[-c(10)]

# #make can_poly wide format
# can_poly_2<-can_poly_cov_all%>%
#     spread(key=class, value=proportion)%>%
#     rename("canopy_shrub_dec_in"="Shrub-Dec", "canopy_tree_con_in"="Tree-Con", "canopy_tree_dec_in"="Tree-Dec", "canopy_grass_in"="Grass", "canopy_shrub_con_in"="Shrub-Con")%>%
#     select(-c(12))


#make UPLVI wide format
UPLVI_PRIMECLAS<-spread(UPLVI_class_cov_all_in$PRIMECLAS, class, proportion)%>%
  dplyr::select(-c(6))
  colnames(UPLVI_PRIMECLAS)[7:8] <- paste("PRIMECLASS_", colnames(UPLVI_PRIMECLAS[,c(7:8)]), sep = "")
UPLVI_LANDCLAS<-spread(UPLVI_class_cov_all_in$LANDCLAS, class, proportion)%>%
  dplyr::select(-c(6))
  colnames(UPLVI_LANDCLAS)[7:12] <- paste("LANDCLAS_", colnames(UPLVI_LANDCLAS[,c(7:12)]), sep = "")
UPLVI_STYPE<-spread(UPLVI_class_cov_all_in$STYPE, class, proportion)%>%
  dplyr::select(-c(6))
  colnames(UPLVI_STYPE)[7:34] <- paste("STYPE_", colnames(UPLVI_STYPE[,c(7:34)]), sep = "")


pp_all_Met_wide_in<-pp_ext_r_cent_in%>%
  left_join(prop_Trail_all[c(2, 8)])%>%
  rename("trail_proportion"="proportion")%>%
  left_join(as.data.frame(tot_trail_in[c(2,6)]))%>%
  dplyr::select(-(16))%>%
  left_join(prop_Road_all[c(2,8)])%>%
  rename("road_proportion" ="proportion")%>%
  left_join(as.data.frame(tot_road_in[c(2,7)]))%>%
  dplyr::select(-(18))%>%
  #left_join(can_poly_2[c(2, 7:11 )])%>%
  left_join(UPLVI_PRIMECLAS[c(2, 7:8)])%>%
  left_join(UPLVI_LANDCLAS[c(2, 7:12)])%>%
  left_join(UPLVI_STYPE[c(2, 7:34)])

#replace NA's with 0



save(pp_all_Met_wide_in , file="3_pipeline/store/pp_all_Met_wide_in.rData")

```



#### long format (for shiny) {-}
```{r}
load("3_pipeline/store/prop_Road_all.rData")
load("3_pipeline/store/prop_Trail_all.rData")
load("3_pipeline/store/UPLVI_class_cov_all.rData")
load("3_pipeline/store/can_poly_cov_all.rData")

library(units)
can_poly_cov_all$pp_area<-as_units(can_poly_cov_all$pp_area, "m2")

#add metrics to comnmon list
pp_ext_metrics_all<-c(list(prop_Road_all, prop_Trail_all, can_poly_cov_all), UPLVI_class_cov_all)
names(pp_ext_metrics_all) <- c("road_prop", "trail_prop", "canopy_class", "PRIMECLAS", "LANDCLAS", "STYPE")

save(pp_ext_metrics_all, file="3_pipeline/store/pp_ext_metrics_all.rData")

#make a data.frame
pp_ext_metrics_all_df<-rbind(prop_Road_all, prop_Trail_all, can_poly_cov_all, UPLVI_class_cov_all$PRIMECLAS, UPLVI_class_cov_all$LANDCLAS, UPLVI_class_cov_all$STYPE)

#add cluster # column
pp_ext_metrics_all_df<-left_join(pp_ext_metrics_all_df, pp_all_clus_RoG_wgs[c(2,6)])
pp_ext_metrics_all_df<-pp_ext_metrics_all_df[-c(10)]

save(pp_ext_metrics_all_df , file="3_pipeline/store/pp_ext_metrics_all_df.rData")

```


## Format sf object for plotting

```{r}
load("1_data/manual/pp_all_RoG.rdata")
pp_all_clus_RoG$cluster<-pp_all_clus_RoG$group
pp_all_clus_RoG$group<- paste(pp_all_clus_RoG$ID, pp_all_clus_RoG$Reach, sep = "_")

#convert to wgs so it works in leaflet
pp_all_clus_RoG_wgs<-st_transform(pp_all_clus_RoG, crs ="+proj=longlat +datum=WGS84")

#caclulate centroid and add the xy cooridinates. This will help plot labels in leaflet
centers <- data.frame(gCentroid(as(pp_all_clus_RoG_wgs, "Spatial"),  byid = TRUE))
pp_all_clus_RoG_wgs<-cbind(pp_all_clus_RoG_wgs, centers)

save(pp_all_clus_RoG_wgs, file="1_data/manual/pp_all_clus_RoG_wgs.rdata")

#do the same for RoG layer
RoG_Reaches<-st_read("1_data/external/Ribbon_of_Green_Study_Area_Reaches/Reach_Boundaries/RoG_Reaches_V1_210301/RoG_Reaches_V1_210301.shp")
RoG_Reaches_wgs<-st_transform(RoG_Reaches, crs ="+proj=longlat +datum=WGS84")
centers <- data.frame(gCentroid(as(RoG_Reaches_wgs, "Spatial"),  byid = TRUE))
RoG_Reaches_wgs<-cbind(RoG_Reaches_wgs, centers)

save(RoG_Reaches_wgs, file="1_data/manual/RoG_Reaches_wgs.rdata")
```










```{r eval=FALSE}
can_poly_cov_all_2<-
    can_poly_cov_all%>% 
    dplyr::select(-c(3)) %>% 
    spread(key=Class, value=proportion)%>%
    rename("Shrub_Con"="Shrub-Con", "Shrub_Dec"="Shrub-Dec", "Tree_Con"="Tree-Con", "Tree_Dec"="Tree-Dec", "ID"="group")

UPLVI_PRIMECLAS<-spread(UPLVI_class_cov_all$PRIMECLAS[-c(3:5, 7)], PRIMECLAS, t_proportion)
UPLVI_LANDCLAS<-spread(UPLVI_class_cov_all$LANDCLAS[-c(3:5, 7)], LANDCLAS, t_proportion)
UPLVI_STYPE<-spread(UPLVI_class_cov_all$STYPE[-c(3:5, 7)], STYPE, t_proportion)


prop_Road_all<-prop_Road_all %>% 
  rename("r_proportion"="proportion") %>% 
  dplyr::select(-c("pp_area"))


prop_Trail_all<- prop_Trail_all %>% 
  rename("t_proportion"="proportion") %>% 
  dplyr::select(-c("pp_area"))

prop_Road_all<-prop_Road_all%>%
  rename("ID"="group")

prop_Trail_all<-prop_Trail_all%>%
  rename("ID"="group")

# trail_tot<-as.data.frame(trail_tot)%>%
#   rename("ID"="group")%>%
#   dplyr::select(-c(5:7))
# 
# road_tot<-as.data.frame(road_tot)%>%
#   rename("ID"="group")%>%
#   dplyr::select(-c(5:7))

pp_metrics<-pp_all_clus_RoG%>%
  left_join(can_poly_cov_all_2, by=c("group"="ID"))%>%
  left_join(UPLVI_PRIMECLAS, by="group")%>%
  left_join(UPLVI_LANDCLAS, by="group")%>%
  left_join(UPLVI_STYPE, by="group")%>%
  left_join(prop_Road_all, by=c("group"="ID"))%>%
  #left_join(road_tot)%>%
  left_join(prop_Trail_all, by=c("group"="ID"))
  #left_join(trail_tot)

#remove unneeded columns
pp_metrics<-pp_metrics[-c(11,14,21)]
pp_metrics<-rename(pp_metrics, "ID"="ID.x")
pp_all_metrics<-pp_metrics

save(pp_all_metrics, file="1_data/manual/pp_all_metrics.rData")

# change crs to wgs for shiny app
pp_all_metrics_wgs<-st_transform(pp_all_metrics, crs ="+proj=longlat +datum=WGS84")
save(pp_all_metrics_wgs, file="1_data/manual/pp_all_metrics_wgs.rdata")


# filter out polygons based on minimum area
## set min area (m^2)
m<-5000

library(units) #use this package to convert the threshold value to a units object
m=set_units(m, m^2)
f<-pp_all_metrics_wgs$area>m
pp_all_metrics_wgs_fil=pp_all_metrics_wgs[f,]

save(pp_all_metrics_wgs_fil, file="3_pipeline/store/pp_all_metrics_wgs_fil.rData")

```


